Impact of grids and dynamical cores in CESM2.2 on the surface mass balance of the Greenland Ice Sheet

### 2022-09-10 first_review/response-to-reviewers.tex

%Cullather, R. I., Nowicki, S. M. J., Zhao, B., & Suarez, M. J. (2014) Evaluation of the surface representation of the Greenland Ice Sheet in a General Circulation Model. J. Climate, 27 (13), 4835-4856. \newline

%Helsen, M. M., van de Wal, R. S. W., Reerink, T. S., Bitanja, R., Madsen, M. S., Yang, S., Li, Q., & Zhang, Q. (2017) On the importance of the albedo parameterization for the mass balance of the Greenland ice sheet in EC-Earth. The Cryosphere, 11, 1949-2017. \newline

%Alexander, P. M., LeGrande, A. N., Fischer, E., Tedesco, M., Fettweis, X., Kelley, M., Nowicki, S. M. J., & Schmidt, G. A. (2019) Simulated Greenland surface mass balance in the GISS ModelE2 GCM: Role of the ice sheet surface. J. Geophys. Res., 124 (3), 750- 765.

{\color{blue}{We thank the reviewer for this point. We had not thought too much about albedo and so it's for the best that this was brought up. This is an interesting hypothesis, that in the coarser grids, the negative precip biases along the ice sheet margins may lead to more frequent bare ice exposure, and through it's lesser albedo, more melt. We have explored whether there is any evidence for this in our simulations, and do find that there is a negative albedo anomaly around the southeast, southwest and northwest coasts in the coarse grids. However, this is only with respect to the RACMO simulations.}}

@@

4. Line 94: Why is “conversely” used here? It seems like a word such as “Additionally” would be more appropriate.  \newline

{\color{blue}{Sentence has been reworded.}} \newline

5. Lines 115-116: Clarify that this is a positive precipitation bias in the interior.  \newline

6. Lines 111-123: Here there is no mention of elevation classes. I suggest mentioning them briefly here as this is a way to increase SMB resolution (from the standpoint of runoff but not precipitation) that is used in CESM  \newline

{\color{blue}{We agree that the idea of ECs should be introduced here, and have added a sentence.}} \newline

7. Figure 1, caption: Please provide a brief description of each of the grids here for the benefit of the reader. E.g. (a) 2 latitude-longitude grid (f19), etc.  \newline

@@

8. Line 189: Briefly explain “cubed-sphere”.  \newline

{\color{blue}{Done.}} \newline

9. Table 1: It would be helpful to specify which grids are latitude-longitude and which are unstructured either as a column in the table or text in the caption.  \newline

{\color{blue}{Added to the text which grids are lat-lon and which are unstructured.}} \newline

@@

10. Lines 229-230: Does ne30 indicate 30 quadrilateral elements per face and pg3 indicate 3 control volumes per element? Please clarify. \newline

{\color{blue}{The text indicates that ne30 indicates are 30x30 elements per cube-face, and pg3 refers to 3x3 control volumes per element. The global number of grid columns is then 6 faces * 30x30 elements per face * 3x3 control volumes per element.}} \newline

@@

12. Line 249: This is the first time RACMO is mentioned. Please explain briefly what RACMO is here.  \newline

{\color{blue}{Fixed.}} \newline

13. Line 254: Clarify that “expensive” means “computationally expensive.” \newline

{\color{blue}{This information has been absorbed into a new sub-section on computational costs. References to ``cheaper" or ``more expensive" are more obviously referring computational costs as that is the title of the section.}} \newline

@@

14. Line 301: Though not essential it would be interesting to know the lapse rate that is assumed. \newline

{\color{blue}{6.5K/km. I'd refer to reviewer to the Sellevold et al paper referenced, for more details on the choice of lapse rate.}} \newline

@@

15. Line 314: I’m confused as to why accumulation below 10 m does not contribute to SMB. How is this addition of mass accounted for? \newline

{\color{blue}{The internally computed SMB is a bit strange, in that only tracks ice. It does not consider the overlying 10m of snow to contribute to the SMB. In response to the other reviewer, we've clarified this in sections 2.5.}} \newline

@@

16. Lines 318-319: What is the period for the prescribed SSTs? \newline

{\color{blue}{Perpetual 1979. Added this in parantheses.}} \newline

@@

17. Lines 322-338: This section only includes discussion of RACMO and in-situ measurements, which are not used, while ERA5, CERES and CALIPSO are not at all discussed. These validation products must be documented here as well. \newline

{\color{blue}{Fixed.}} \newline

18. Line 322: RACMO2 is still a model, and does not fall under an “observational” dataset. I suggest changing this to read “Validation Datasets”, “Evaluation Datasets” or “Comparison Datasets”. \newline

{\color{blue}{Changed to validation dataset throughout the manuscript.}} \newline

@@

19. Figure 3 caption: I believe the left and right figures are switched here. Also, the fact that RACMO RCM simulations are included should be mentioned. \newline

{\color{blue}{Fixed, and included references to the RACMO grids in the text and figure caption.}} \newline

@@

20. Formula 1: The f’s in runoff seem to have been turned into function symbols. Also I suggest using “evaporation + sublimation” to avoid confusion with division. \newline

{\color{blue}{Switched to type writer font. Used reviewers suggestion to replace slash with a plus sign.}} \newline

21. Line 406: I believe this should read “expressed as temperature differences”.

{\color{blue}{Did something similar, as temperature was missing from the sentence.}} \newline

22. Lines 408-416, 429-433: How can the effect be primarily ascribed to condensational heating if the temperature differences do not align in magnitude or spatial extent? Are there feedbacks involved? Please clarify in the text, or qualify the statements. \newline

{\color{blue}{The imperfect alignment of the clubb heating anomalies and the temperature anomalies has a lot to do with averaging the clubb heating over all time. If you look at Figure 10 in Herrington and Reed QJRMS paper, the alignment is strongest when you sample the clubb heating when/where the dycore has upward vertical velocities. Whereas there is no correspondence with the temperature anomalies for descending grid columns. So since our clubb heating is just a climatology, we lose that alignment with the temperature anomalies. But it's clear that a strong heating signal is piercing through in the climatology anyway, and so I'm strongly implying that this is consistent with QJRMS mechanism. I've added an entire paragraph discussing this misalignment.}} \newline

@@

23. Lines 437-438: This also seems to be true in other cases. \newline

{\color{blue}{See above comment.}} \newline

24. Lines 440-441: This seems consistent with other results that increasing resolution led to a temperature increase. \newline

{\color{blue}{This response appears to be different from the other cases; with clear changes to the planetary waves that are not seen in the other simulations.}} \newline

@@

26. Line 479: Clarify that cooler temperatures in the coarse resolution simulations are being referred to here. \newline

{\color{blue}{Decided to leave it as.}} \newline

27. Figure 8: The CALIPSO grid appears to be coarser resolution than the CERES grid. Is this a 2 grid? \newline

{\color{blue}{Yes, CALIPSO is on a 2 degree grid. Corrected Table 2. Great catch!}} \newline

@@

28. Lines 573-575: Wouldn’t the elevation class scheme fix this problem by raising the ice-covered sub-grid portion to a higher elevation? \newline

{\color{blue}{That's a good point that we overlooked. We have added a sentence to the end of the next paragraph clarifying that differences between the grid cell mean and actual ice surface are not necessarily problematic due to the EC scheme.}} \newline

@@

29. Lines 606-616: If the biases are larger at low elevations, it could also be coincidental that these areas also have lower ice fractions. I think the authors should also consider the possibility of other factors. \newline

{\color{blue}{We've added references to this coincidence in response to your related questions in the general remarks.}} \newline

@@

34. Line 731: Add “is the” before “surface geopotential”. \newline

{\color{blue}{Done.}} \newline

35. Equation A5: How is T(ref) computed? \newline

{\color{blue}{Defined in equation A1.}} \newline

36. Lines 740-748: Here some terms should be defined, including Tv, qv and p0. \newline

{\color{blue}{Added definitions for Tv and thetav. p0 is defined just below equation A3. I don't see a ``qv".}}

### 2022-09-10 second_submit/arcticgrids.tex
updates

Six different configurations, a mixture of grids and atmospheric dynamical cores available in the Community Earth System Model, version 2.2 (CESM2.2), are evaluated for their skill in representing the climate of the Arctic and the surface mass balance of the Greenland Ice Sheet (GrIS). The conventional $1^{\circ}-2^{\circ}$ resolution grids systematically overestimate both accumulation and ablation over the GrIS. Of these conventional grids, the latitude-longitude grids outperform the quasi-uniform unstructured grids because they have more degrees of freedom to represent the GrIS. Two Arctic-refined meshes, with $1/4^{\circ}$ and $1/8^{\circ}$ refinement over Greenland, are documented as newly supported configurations in CESM2.2. The Arctic meshes substantially improve the simulated clouds and precipitation rates in the Arctic. Over Greenland, these meshes skillfully represent accumulation and ablation processes, leading to a more realistic GrIS surface mass balance. As CESM is in the process of transitioning away from conventional latitude-longitude grids, these new Arctic-refined meshes improve the representation of polar processes in CESM by recovering resolution lost in the transition to quasi-uniform grids, albeit at increased computational cost.

@@

General Circulation Models (GCMs) are powerful tools for understanding the meteorology and climate of the high latitudes, which are among the most sensitive regions on Earth to global and environmental change. GCMs differ vastly in their numerical treatment of polar regions because of the so-called \textit{pole problem} \cite{W2007JMSJ}. The pole problem refers to numerical instability arising from the convergence of meridian lines into polar singularities on latitude-longitude grids (e.g., Figure~\ref{fig:uni-grids}a, hereafter referred to as \textit{lat-lon} grids). Depending on the numerics, methods exist to suppress this instability, and lat-lon grids may be advantageous for polar processes by representing structures with finer resolution than elsewhere in the computational domain. With the recent trend towards quasi-uniform unstructured grids, any potential benefits of lat-lon grids in polar regions may be lost {\color{blue}{(hereafter, {\textit{quasi-uniform}} refers to approximately isotropic grids, in contrast to lat-lon grids, which are highly anisotropic)}}. In this study, we evaluate a number of grids and dynamical cores (hereafter referred to as \textit{dycores}) available in the Community Earth System Model, version 2.2 \cite<CESM2.2;>{CESM2}, including new variable-resolution grids, to understand their impacts on the simulated Arctic climate. We focus specifically on the climate and surface mass balance of the Greenland Ice Sheet.

@@

The SMB of the Greenland Ice Sheet (hereafter \textit{GrIS}) is determined by processes occurring over a range of scales that are difficult to represent in GCMs \cite{P2010CC}. GrIS precipitation is concentrated at the ice-sheet margins, where steep topographic slopes drive orographic precipitation. The truncated topography used by low resolution GCMs enables moisture to penetrate well into the GrIS interior, manifesting as a positive precipitation bias \cite{P2000GPC,VETAL2018TC}. GrIS ablation areas {\color{blue}{(marginal regions where the annual SMB is negative)}} are typically less than 100\,km wide and are confined to low-lying areas or regions with low precipitation. These narrow ablation zones are not fully resolved in low-resolution GCMs, and may further degrade the simulated SMB. For example, CESM, version 2.0 (CESM2) underestimates ablation in the northern GrIS, leading to unrealistic ice advance when run with an interactive ice sheet component \cite{LETAL2020JAMES}.

@@

\begin{figure}[t]
\begin{center}
%\begin{tabular}{cc}
         \includegraphics[width=100mm]{grid-uniform}
         %\includegraphics[width=50mm]{grid-f19.pdf}&
         %\includegraphics[width=50mm]{grid-f09.pdf}\\
         %\includegraphics[width=50mm]{grid-ne30pg2.pdf}&
         %\includegraphics[width=50mm]{grid-ne30pg3.pdf} \\
%\end{tabular}
\end{center}
\caption{Computational grids for the $1^{\circ}-2^{\circ}$ {\color{blue}{lat-lon and quasi-uniform unstructured}} grids in this study. Grids names after Table~\ref{tbl:table1}, (a) $\tt f19$, (b) $\tt f09$, (c) $\tt ne30pg2$ and (d) $\tt ne30pg3$.}
\label{fig:uni-grids}
\end{figure}

\begin{figure}[t]
\begin{center}
%\begin{tabular}{ccc}
         \includegraphics[width=140mm]{grid-varres.pdf}
         %\includegraphics[width=40mm]{grid-ARCTIC.pdf}&
         %\includegraphics[width=40mm]{grid-ARCTICGRIS.pdf}&
         %\includegraphics[width=40mm]{grid-CONUS.pdf} \\
%\end{tabular}
\end{center}
\caption{Variable-resolution grids available in CESM2.2, (a) $\tt Arctic$, (b) $\tt Arctic-GrIS$ and (c) $\tt CONUS$. Note what is shown is the element grid; the computational grid has $3\times3$ independent grid points per element.}
\label{fig:vr-grids}
\end{figure}

@@

The SE dycore integrates the {\color{blue}{hydrostatic}} equations of motion using a high-order continuous Galerkin method \cite{TTI1997JCP,TF2010JCP}. The computational domain is a cubed-sphere grid tiled with quadrilateral elements (see Figure~\ref{fig:vr-grids}). Each element contains a fourth-order basis set in each horizontal direction, with the solution defined at the roots of the basis functions, the Gauss-Lobatto-Legendre quadrature points. This results in 16 nodal points per element, with 12 of the points lying on the (shared) element boundary. Communication between elements uses the direct stiffness summation \cite{canuto2007}, which applies a numerical flux to the element boundaries to reconcile overlapping nodal values and produce a continuous global basis set.

@@

Three VR meshes were developed for the CESM2.2 release to support grid refinement over the Arctic and the United States (Figure~\ref{fig:vr-grids}). This paper serves as the official documentation of these grids. The VR meshes were developed using the software package SQuadgen (\url{https://github.com/ClimateGlobalChange/squadgen}). The $\tt Arctic$ grid is a $1^{\circ}$ grid with $1/4^{\circ}$ regional refinement over the broader Arctic region. The $\tt Arctic-GrIS$ grid is identical to the $\tt Arctic$ grid, but with an additional patch covering the island of Greenland with $1/8^{\circ}$ resolution. The $\tt CONUS$ grid contains $1/8^{\circ}$ refinement over the United States, and $1^{\circ}$ everywhere else. The $\tt CONUS$ grid is not discussed any further in this paper; see \citeA{PETAL2020BAMS} for simulations with the $\tt CONUS$ grid.

 \begin{table*}
 \centering
 \scriptsize
 \begin{tabular}{lcccccc}
   \hline
   grid name & dycore & $\Delta x_{\mathrm{eq}}$ (km) & $\Delta x_{\mathrm{fine}}$ (km) & $\Delta t_{\mathrm{phys}}$ (s) & {\color{blue}{cost(25 nodes)}} & {\color{blue}{cost(50 nodes)}} \\ 
   \hline
   $\tt f19$ & FV & 278 & - &1800 & {\color{blue}{436.66}} & {\color{blue}{-}} \\
   $\tt f09$ & FV & 139 & - &1800 & {\color{blue}{1534.57}} & {\color{blue}{2024.24}} \\
   $\tt ne30pg2$ & SE-CSLAM & 167 & - & 1800 & {\color{blue}{1497.26}} & {\color{blue}{1683.97}} \\
   $\tt ne30pg3$ & SE-CSLAM & 111 & - & 1800 & {\color{blue}{1890.48}} & {\color{blue}{2090.43}} \\
   $\tt ne30pg3^{*}$ & SE-CSLAM & 111 & - & 450 & {\color{blue}{-}} & {\color{blue}{-}} \\
   $\tt Arctic$ & SE & 111 & 28 & 450 & {\color{blue}{15947.41}} & {\color{blue}{16675.45}} \\
   $\tt Arctic-GrIS$ & SE & 111 & 14 & 225 & {\color{blue}{40305.03}} & {\color{blue}{41036.67}} \\
 \hline
 \end{tabular}
  \caption{Grids and dycores used in this study. $\Delta x_{\mathrm{eq}}$ is the average equatorial grid spacing, $\Delta x_{\mathrm{fine}}$ is the grid spacing in the refined region (if applicable), and $\Delta t_{\mathrm{phys}}$ is the physics time step. FV refers to the finite-volume dycore, SE the spectral-element dycore, and SE-CSLAM the spectral-element dycore with CSLAM tracer advection.  We use the $\tt ne30pg3$ grid for two runs with different values of $\Delta t_{\mathrm{phys}}$. {\color{blue}{The last columns provide the computational costs in core hours per simulated year (CHPSY). The costs are from single month runs using 25 nodes and 50 nodes on the Cheyenne supercomputer \cite{cheyenne}.}}}
 \label{tbl:table1}
 \end{table*}

@@

Since snow in the accumulation zone must reach the cap to simulate a positive SMB, the snow depths on the VR grids were spun up by forcing CLM in standalone mode, cycling over data from a 20-year $\tt Arctic$ $\tt FHIST$ simulation for about 500 years. The $1^{\circ}-2^{\circ}$ lat-lon and quasi-uniform unstructured grids are initialized with the SMB from an existing $\tt f09$ spun-up initial condition.

@@

 \begin{table*}
 \centering
 \scriptsize
 \begin{tabular}{lcccc}
   \hline
   data product & years used in this study & resolution & citation \\ 
   \hline
   ERA5 & 1979-1998 & $1/4^{\circ}$ & \citeA{ERA5} \\
   CERES-EBAF ED4.1 & 2003-2020 & $1^{\circ}$ & \citeA{CERES-EBAF} \\
   CALIPSO-GOCCP & 2006-2017 & $1^{\circ}$ & \citeA{CALIPSO-GOCCP} \\
   RACMO2.3 & 1979-1998 & 11 km & \citeA{NETAL2015TC} \\
   RACMO2.3p2 & 1979-1998 & 5.5 km & \citeA{NETAL2019SCIENCE} \\
 \hline
 \end{tabular}
  \caption{Description of {\color{blue}{validation}} datasets used in this study.}
 \label{tbl:table2}
 \end{table*}

@@

\begin{figure}[t]
\begin{center}
         \includegraphics[width=100mm]{temp_grisres.pdf}
\end{center}
\caption{The spatial properties of the GrIS as represented by different grids in this study. (Left) GrIS area error, computed as the relative differences from a 4-km dataset used to create the CESM ice masks, (right) approximate average grid spacing over GrIS.}
\label{fig:grisdx}
\end{figure}

@@

The SMB is expressed in a form that is agnostic of water phase, a total water mass balance, to facilitate comparisons across different grids with different ice masks and to increase consistency with the variables available in the RACMO datasets. The SMB for total water can be expressed as:
\begin{equation}
SMB = accumulation + runoff + evaporation + sublimation, \label{eq:SMB}
\end{equation}
where all terms have consistent sign conventions (positive values contribute mass, and negative values reduce mass). The accumulation source term refers to the combined solid and liquid precipitation, runoff refers to the liquid water sink, and evaporation/sublimation is the vapor sink. Since the runoff term aggregates many processes, we isolate the melting contribution by also tracking the combined melt of snow and ice.

@@

Before delving into the simulated Arctic climate conditions, we assess the global mean differences between the various grids and dycores. Figure~\ref{fig:dT-lores} shows 1979-1998 annual mean, zonal mean height plots expressed as differences between {\color{blue}{$1^{\circ}-2^{\circ}$ lat-lon and quasi-uniform unstructured}} grids and dycores. The $\tt f09$ grid is warmer than the $\tt f19$ grid, primarily in the mid-to-high latitudes throughout the depth of the troposphere. This is a common response to increasing horizontal resolution in GCMs \cite{PS2002CD,RETAL2006JC}.
\citeA{HK2020QJRMS} have shown that this occurs in CAM due to higher resolved vertical velocities which, in turn, generate more condensational heating in the CLUBB macrophyiscs. The right panel in Figure~\ref{fig:dT-lores}a supports this interpretation, showing an increase in the climatological CLUBB heating at all latitudes in the $\tt f09$ grid, but with the largest increase in the mid-latitudes.

@@

@@

The transects in Figure~\ref{fig:ztrans} show that the ice sheet surface on the coarse grids is not systematically lower than the true surface in ablation zones. Rather, the smoothing and flattening of the raw topography, necessary to prevent the model from exciting grid-scale numerical modes, causes the lower-elevation ablation zones to extend beyond the true ice sheet margin, causing the modeled ablation zones (which must reside within the ice sheet mask) to be elevated relative to the actual ice surface. The $\tt f19$ grid has both the smoothest topography and the flattest ice sheet since its dynamics are coarsest, whereas the $\tt f09$, $\tt ne30pg2$ and $\tt ne30pg3$ grids have similar dynamical resolution and use identical smoothing. This suggests that coarser models will tend to elevate the ablation zones {\color{blue}{relative to where they should be, which may be expected to cause anomalous (adiabatic) cooling and}} depressed melt rates, but this is opposite the melt bias that occurs in the coarse grid simulations.

>>>>

Six different configurations, a mixture of grids and atmospheric dynamical cores available in the Community Earth System Model, version 2.2 (CESM2.2), are evaluated for their skill in representing the climate of the Arctic and the surface mass balance of the Greenland Ice Sheet (GrIS). {\color{blue}{The finite-volume dynamical core uses structured, latitude-longitude grids, whereas the spectral-element dynamical core is built on unstructured meshes, permitting grid flexibility such as quasi-uniform grid spacing globally.}} The $1^{\circ}-2^{\circ}$ {\color{blue}{latitude-longitude and quasi-uniform unstructured}} grids systematically overestimate both accumulation and ablation over the GrIS. Of these $1^{\circ}-2^{\circ}$ grids, the latitude-longitude grids outperform the quasi-uniform unstructured grids because they have more degrees of freedom to represent the GrIS. Two Arctic-refined meshes, with $1/4^{\circ}$ and $1/8^{\circ}$ refinement over Greenland, were developed for the spectral-element dynamical core and are documented here as newly supported configurations in CESM2.2. The Arctic meshes substantially improve the simulated clouds and precipitation rates in the Arctic. Over Greenland, these meshes skillfully represent accumulation and ablation processes, leading to a more realistic GrIS surface mass balance. As CESM is in the process of transitioning away from conventional latitude-longitude grids, these new Arctic-refined meshes improve the representation of polar processes in CESM by recovering resolution lost in the transition to quasi-uniform grids, albeit at increased computational cost.

@@

General Circulation Models (GCMs) are powerful tools for understanding the meteorology and climate of the high latitudes, which are among the most sensitive regions on Earth to global and environmental change. GCMs differ vastly in their numerical treatment of polar regions because of the so-called \textit{pole problem} \cite{W2007JMSJ}. The pole problem refers to numerical instability arising from the convergence of meridian lines into polar singularities on latitude-longitude grids (e.g., Figure~\ref{fig:uni-grids}a, hereafter referred to as \textit{lat-lon} grids). Depending on the numerics, methods exist to suppress this instability, and lat-lon grids may be advantageous for polar processes by representing structures with finer resolution than elsewhere in the computational domain. With the recent trend towards quasi-uniform unstructured grids, any potential benefits of lat-lon grids in polar regions may be lost {\color{blue}{(hereafter, {\textit{quasi-uniform}} refers to approximately isotropic grids, in contrast to lat-lon grids, which are highly anisotropic due to the polar singularity)}}. In this study, we evaluate a number of grids and dynamical cores (hereafter referred to as \textit{dycores}) available in the Community Earth System Model, version 2.2 \cite<CESM2.2;>{CESM2}, including new variable-resolution grids {\color{blue}{(i.e., grids with enhanced resolution over a particular region)}}, to understand their impacts on the simulated Arctic climate. We focus specifically on the climate and surface mass balance of the Greenland Ice Sheet.

@@

The SMB of the Greenland Ice Sheet (hereafter \textit{GrIS}) is determined by processes occurring over a range of scales that are difficult to represent in GCMs \cite{P2010CC}. GrIS precipitation is concentrated at the ice-sheet margins, where steep topographic slopes drive orographic precipitation. The truncated topography used by low resolution GCMs enables moisture to penetrate well into the GrIS interior, manifesting as a positive precipitation bias {\color{blue}{in the interior}} \cite{P2000GPC,VETAL2018TC}. GrIS ablation areas {\color{blue}{(marginal regions where the annual SMB is negative)}} are typically less than 100\,km wide and are confined to low-lying areas or regions with low precipitation. These narrow ablation zones are not fully resolved in low-resolution GCMs, and may further degrade the simulated SMB. {\color{blue}{More recently, GCMs such as CESM and the NASA Goddard Institute for Space Studies GCM \cite{AETAL2019JGR}, have implemented an elevation class downscaling scheme for computing the SMB. The downscaling helps to resolve to these narrow ablation zones in GCMs \cite{SETAL2019TC}, but large SMB biases remain.}} For example, CESM, version 2.0 (CESM2) underestimates ablation in the northern GrIS, leading to unrealistic ice advance when run with an interactive ice sheet component \cite{LETAL2020JAMES}.

@@

\begin{figure}[t]
\begin{center}
         \includegraphics[width=100mm]{grid-uniform}
\end{center}
\caption{Computational grids for the $1^{\circ}-2^{\circ}$ lat-lon and quasi-uniform unstructured grids in this study. Grids names after Table~\ref{tbl:table1}, (a) $2^{\circ}$ lat-lon grid ($\tt f19$), (b) $1^{\circ}$ lat-lon grid ($\tt f09$), (c) $1^{\circ}$ quasi-uniform unstructured grid with reduced physics resolution ($\tt ne30pg2$) and (d) $1^{\circ}$ quasi-uniform unstructured grid ($\tt ne30pg3$).}
\label{fig:uni-grids}
\end{figure}

\begin{figure}[t]
\begin{center}
         \includegraphics[width=140mm]{grid-varres.pdf}
\end{center}
\caption{Variable-resolution grids available in CESM2.2, (a) $\tt Arctic$, (b) $\tt Arctic-GrIS$ and (c) $\tt CONUS$. Note what is shown is the element grid; the computational grid has $3\times3$ independent grid points per element.}
\label{fig:vr-grids}
\end{figure}

@@

The SE dycore integrates the {\color{blue}{hydrostatic}} equations of motion using a high-order continuous Galerkin method \cite{TTI1997JCP,TF2010JCP}. The computational domain is a cubed-sphere grid ({\color{blue}{A cubed-sphere grid is obtained by projecting each face of a cube onto a sphere}}) tiled with quadrilateral elements (see Figure~\ref{fig:vr-grids}). Each element contains a fourth-order basis set in each horizontal direction, with the solution defined at the roots of the basis functions, the Gauss-Lobatto-Legendre quadrature points. This results in 16 nodal points per element, with 12 of the points lying on the (shared) element boundary. Communication between elements uses the direct stiffness summation \cite{canuto2007}, which applies a numerical flux to the element boundaries to reconcile overlapping nodal values and produce a continuous global basis set.

@@

Three VR meshes were developed for the CESM2.2 release to support grid refinement over the Arctic and the United States (Figure~\ref{fig:vr-grids}). This paper serves as the official documentation of these grids. The VR meshes were developed using the software package SQuadgen (\url{https://github.com/ClimateGlobalChange/squadgen}). The $\tt Arctic$ grid is a $1^{\circ}$ grid with $1/4^{\circ}$ regional refinement over the broader Arctic region ({\color{blue}{when using degrees to describe the resolution of unstructured grids, degrees refers to degrees at the equator}}). The $\tt Arctic-GrIS$ grid is identical to the $\tt Arctic$ grid, but with an additional patch covering the island of Greenland with $1/8^{\circ}$ resolution. The $\tt CONUS$ grid contains $1/8^{\circ}$ refinement over the United States, and $1^{\circ}$ everywhere else. The $\tt CONUS$ grid is not discussed any further in this paper; see \citeA{PETAL2020BAMS} for simulations with the $\tt CONUS$ grid.

 \begin{table*}
 \centering
 \scriptsize
 \begin{tabular}{lcccccc}
   \hline
   grid name & dycore & $\Delta x_{\mathrm{eq}}$ (km) & $\Delta x_{\mathrm{fine}}$ (km) & $\Delta t_{\mathrm{phys}}$ (s) & {\color{blue}{cost(25 nodes)}} & {\color{blue}{cost(50 nodes)}} \\ 
   \hline
   $\tt f19$ & FV & 278 & - &1800 & {\color{blue}{436.66}} & {\color{blue}{-}} \\
   $\tt f09$ & FV & 139 & - &1800 & {\color{blue}{1534.57}} & {\color{blue}{2024.24}} \\
   $\tt ne30pg2$ & SE-CSLAM & 167 & - & 1800 & {\color{blue}{1497.26}} & {\color{blue}{1683.97}} \\
   $\tt ne30pg3$ & SE-CSLAM & 111 & - & 1800 & {\color{blue}{1890.48}} & {\color{blue}{2090.43}} \\
   $\tt ne30pg3^{*}$ & SE-CSLAM & 111 & - & 450 & {\color{blue}{-}} & {\color{blue}{-}} \\
   $\tt Arctic$ & SE & 111 & 28 & 450 & {\color{blue}{15947.41}} & {\color{blue}{16675.45}} \\
   $\tt Arctic-GrIS$ & SE & 111 & 14 & 225 & {\color{blue}{40305.03}} & {\color{blue}{41036.67}} \\
 \hline
 \end{tabular}
  \caption{Grids and dycores used in this study. $\Delta x_{\mathrm{eq}}$ is the average equatorial grid spacing, $\Delta x_{\mathrm{fine}}$ is the grid spacing in the refined region (if applicable), and $\Delta t_{\mathrm{phys}}$ is the physics time step. FV refers to the finite-volume dycore, SE the spectral-element dycore, and SE-CSLAM the spectral-element dycore with CSLAM tracer advection. {\color{blue}{The FV dycore uses lat-lon grids, whereas the SE and SE-CSLAM dycores run on unstructured grids.}} We use the $\tt ne30pg3$ grid for two runs with different values of $\Delta t_{\mathrm{phys}}$. {\color{blue}{The last columns provide the computational costs in core hours per simulated year (CHPSY). The costs are from single month runs using 25 nodes and 50 nodes on the Cheyenne supercomputer \cite{cheyenne}.}}}
 \label{tbl:table1}
 \end{table*}

@@

Since snow in the accumulation zone must reach the cap to simulate a positive SMB, the snow depths on the VR grids were spun up by forcing CLM in standalone mode, cycling over data from a 20-year $\tt Arctic$ $\tt FHIST$ simulation ({\color{blue}{with perpetual 1979 boundary conditions}}) for about 500 years. The $1^{\circ}-2^{\circ}$ lat-lon and quasi-uniform unstructured grids are initialized with the SMB from an existing $\tt f09$ spun-up initial condition.

@@

 \begin{table*}
 \centering
 \scriptsize
 \begin{tabular}{lcccc}
   \hline
   data product & years used in this study & resolution & citation \\ 
   \hline
   ERA5 & 1979-1998 & $1/4^{\circ}$ & \citeA{ERA5} \\
   CERES-EBAF ED4.1 & 2003-2020 & $1^{\circ}$ & \citeA{CERES-EBAF} \\
   CALIPSO-GOCCP & 2006-2017 & {\color{blue}{$2^{\circ}$}} & \citeA{CALIPSO-GOCCP} \\
   RACMO2.3 & 1979-1998 & 11 km & \citeA{NETAL2015TC} \\
   RACMO2.3p2 & 1979-1998 & 5.5 km & \citeA{NETAL2019SCIENCE} \\
 \hline
 \end{tabular}
  \caption{Description of {\color{blue}{validation}} datasets used in this study.}
 \label{tbl:table2}
 \end{table*}

@@

\begin{figure}[t]
\begin{center}
         \includegraphics[width=100mm]{temp_grisres.pdf}
\end{center}
\caption{The spatial properties of the GrIS as represented by different grids in this study. (Left) approximate average grid spacing over GrIS, (right) GrIS area error, computed as the relative differences from the reference dataset described in the text. {\color{blue}{Also included are the two RACMO grids used for validation (see Table~\ref{tbl:table2})}}.}
\label{fig:grisdx}
\end{figure}

@@

The SMB is expressed in a form that is agnostic of water phase, a total water mass balance, to facilitate comparisons across different grids with different ice masks and to increase consistency with the variables available in the RACMO datasets. The SMB for total water can be expressed as:
\begin{equation}
\tt SMB = accumulation + runoff + evaporation + sublimation, \label{eq:SMB}
\end{equation}
where all terms have consistent sign conventions (positive values contribute mass, and negative values reduce mass). The accumulation source term refers to the combined solid and liquid precipitation, runoff refers to the liquid water sink, and evaporation/sublimation is the vapor sink. Since the runoff term aggregates many processes, we isolate the melting contribution by also tracking the combined melt of snow and ice.

@@

Before delving into the simulated Arctic climate conditions, we assess the global mean differences between the various grids and dycores. Figure~\ref{fig:dT-lores} shows 1979-1998 annual mean, zonal mean height plots {\color{blue}{of temperature}}, expressed as differences between {\color{blue}{$1^{\circ}-2^{\circ}$ lat-lon and quasi-uniform unstructured}} grids and dycores. The $\tt f09$ grid is warmer than the $\tt f19$ grid, primarily in the mid-to-high latitudes throughout the depth of the troposphere. This is a common response to increasing horizontal resolution in GCMs \cite{PS2002CD,RETAL2006JC}. \citeA{HK2020QJRMS} have shown that this occurs in CAM due to higher resolved vertical velocities which, in turn, generate more condensational heating in the CLUBB macrophyiscs. The right panel in Figure~\ref{fig:dT-lores}a supports this interpretation, showing an increase in the climatological CLUBB heating at all latitudes in the $\tt f09$ grid, but with the largest increase in the mid-latitudes.

@@

{\color{blue}{We note that there is not perfect alignment of the CLUBB heating anomalies and the temperature anomalies in Figures~\ref{fig:dT-lores} $\&$ \ref{fig:dT-hires}. This leaves some uncertainty as to whether CLUBB is solely responsible for the temperature anomalies caused by changing grid resolution or dycore, as asserted above. In the deep Tropics, gravity waves are expected to propagate temperature anomalies far from their heating source, and so the misalignment is less troublesome for this region. In the mid-latitudes, we speculate that this misalignment is due to averaging the CLUBB heating over all time. \citeA{HK2020QJRMS} have shown that only the CLUBB heating coinciding with upward resolved vertical velocities tends to align spatially with the temperature response, whereas heating anomalies associated with descending grid columns had no alignment with the temperature response. Therefore the average over all time will dilute the heating signal, and potentially lead to misalignments between the heating rates and temperature anomalies.}}

@@

The transects in Figure~\ref{fig:ztrans} show that the ice sheet surface on the coarse grids is not systematically lower than the true surface in ablation zones. Rather, the smoothing and flattening of the raw topography, necessary to prevent the model from exciting grid-scale numerical modes, causes the lower-elevation ablation zones to extend beyond the true ice sheet margin, causing the modeled ablation zones (which must reside within the ice sheet mask) to be elevated relative to the actual ice surface. The $\tt f19$ grid has both the smoothest topography and the flattest ice sheet since its dynamics are coarsest, whereas the $\tt f09$, $\tt ne30pg2$ and $\tt ne30pg3$ grids have similar dynamical resolution and use identical smoothing. This suggests that coarser models will tend to elevate the ablation zones {\color{blue}{relative to where they should be, which may be expected to cause anomalous (adiabatic) cooling and}} depressed melt rates, but this is opposite the melt bias that occurs in the coarse grid simulations. {\color{blue}{Also, the EC downscaling should be able to correct for situations where model ice surface is elevated relative to the true ice surface, although for very large elevation differences/errors the EC scheme may not be adequate}}.
